Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue

### 2021-06-17 reports/rmarkdown/manuscript/manuscript_template.Rmd
update

---
title: "Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue"
subtitle: A demonstration of Rmarkdown using Herman Bumpus' data
author: |
  Author One^1^,
  Author Two^2^,
  Author Three^1,2^
keywords: "pandoc, r markdown, knitr"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::pdf_document2:
      keep_tex: true
      toc: false
  bookdown::html_document2: default
  bookdown::word_document2: default
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[L]{XXXXXXXXXXX et al.}
- \fancyhead[R]{MANUSCRIPT SHORT TITLE}
- \usepackage{lineno}
- \linenumbers
linestretch: 1.5
link-citations: yes
linkcolor: blue
csl: nature.csl
cslreferences: nature.csl
bibliography: refs.bib
editor_options:
  chunk_output_type: console
---
\footnotetext[1]{University of Nowhere}
\footnotetext[2]{University of Somewhere}
\footnotetext[3]{University of Lalaland}

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(out.width = '100%',
                      dpi=600,
                      fig.width=10,
                      echo = FALSE,
                      # fig.pos = 'p', # Places figures on their own pages
                      eval = TRUE,
                      results = 'hide',
                      warning = FALSE,
                      message = FALSE )

# I usually load my libraries up front to keep things organized
library(bookdown)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyverse)
library(igraph)
library(phyloseq)
library(vegan)
library(ggpubr)
library(rafalib)
library(scales)
library(dendextend)
library(edgeR)
remotes::install_github("czarnewski/niceRplots",force=T)
library(niceRplots)

#color palettes
pal <- c(RColorBrewer::brewer.pal(9,"Set1"),RColorBrewer::brewer.pal(8,"Set2"),
         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
```

@@

```{r, echo = FALSE, eval = TRUE}
# Load reasult from the analysis
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
patient_groups <- factor(setNames(metadata$joint_clustering,metadata$ID))
bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]

@@

```{r, echo = FALSE, eval = TRUE}
#Assign names to patient groups
big_groups <- c(
  g2_L.iners=    "5",
  g3_L.iners2=   "4",
  g1_L.crispatus= "2",
  g5_Gard=        "3",
  g6_Gard.Prev=   "6",
  g7_Gard.Prev2=  "8",
  g4_Gard.Lacto=  "7",
  g8_Patho=       "1"
)

group_annotation <- as.character(patient_groups)
group_annotation <- names(big_groups)[match(group_annotation,big_groups)]
group_annotation <- as.factor(setNames(group_annotation,names(patient_groups)))

metadata$group_annotation <- group_annotation
write.csv2(metadata,"../../../results/metadata_integration.csv",row.names = T)

@@

#Assign names to bacterial communities
BC_groups <- c(
  BC01= "BC01_Bifidobacterium",
  BC02= "BC02_Mycoplasma",
  BC03= "BC03_Lactobacillus.spp",
  BC04= "BC04_Escherichia.sp",
  BC05= "BC05_Other1",
  BC06= "BC06_Other2",
  BC07= "BC07_Corynebacterium.spp",
  BC08= "BC08_Other3",
  BC09= "BC09_Ureaplasma.sp",
  BC10= "BC10_Campylobacter",
  BC11= "BC11_Mobiluncus_BVAB",
  BC12= "BC12_Faecalibacterium",
  BC13= "BC13_Dialister",
  BC14= "BC14_P.timonesis",
  BC15= "BC15_T.scotoductus"
)

@@

par(mar=c(0,0,0,6))
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] ,
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
      vertex.size=10,main="", layout=l,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
legend(par("usr")[2],par("usr")[4],title.adj = 0,
       legend = levels(factor(group_annotation)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)

@@

#######################
# BACTERIAL ABUNDANCE #
#######################
par(mar=c(4,5,1,12))
temp <- t(t(2^x-1)/colSums(2^x-1))
taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
taxa[is.na(taxa)] <- "NA"
taxa <- taxa[rownames(temp),]
temp <- rowsum(temp , taxa$Class)
temp[is.nan(temp)] <- 0
o <- order( rowSums(temp) , decreasing = T)
sample_o <- order( group_annotation )
temp <- temp[o,sample_o]*100
temp <- temp[-grep("NA",rownames(temp)),]
temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )

taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",
         col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] ,names.arg = rep("",ncol(temp)))
legend(x = par("usr")[c(2)],
       y = par("usr")[c(4)],title.adj = 0,
       legend = rownames(temp),xjust = 0,yjust = 1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
       pt.bg = taxa_pal)
#
# points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-1,
#         col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][sample_o],
#         pch=16,cex=.5,xpd=T)
# points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
#         col= c("grey","red")[as.numeric(metadata$HIVstatus)][sample_o],
#         pch=16,cex=.5,xpd=T)
  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order( group_annotation )],
        pch=15,cex=.8,xpd=T)
  text(par("usr")[1],par("usr")[3]-2,labels = "RepeatedPCR",pos=2,cex=.8,xpd=T)

@@

```{r, echo = FALSE, eval = TRUE,fig.asp=.6}
# Define Layout and set labels
layout(matrix(c(2 ,3,4,5,      6,7,17,17,          17,17,17,17,
                1,1,1,1,      1,8,17,17,          17,17,17,17,
                1,1,1,1,      1,9,18,18,         18,18,18,18,
                1,1,1,1,      1,10,19,19,         19,19,19,19,
                16,15,14,13,  12,11,20,20,         20,20,20,20 ) , ncol= 12,byrow = T),widths = 1,heights = c(1.2,1.2,1.2,1.2,1.2))
# layout.show(12)
figlabels <- letters
#
# m <- matrix(as.character(c(1,2,3,4)),nrow = 2,ncol = 2)
# m
# m2 <- matrix(as.character(c(1,2,3,"m")),nrow = 2,ncol = 2)
# m2
#
# matrix(c(1,2,j,3),nrow = 2,ncol = 2)
# layout_matrix <- function(data, nrow, ncol){
#   return( matrix(as.character(data), nrow = nrow, ncol = ncol)  )
# }
# layout_matrix2 <- function(data, nrow, ncol, widths=c(1,2), heights=c(2,3), res=res){
#   temp <- matrix(as.character(data), nrow = nrow, ncol = ncol)
#   widths <- rep(widths,ncol(temp))[1:ncol(temp)]
#   heights <- rep(heights,nrow(temp))[1:nrow(temp)]
#   
#   cumsum(widths)*100/sum(widths)
#   cumsum(heights)*100/sum(heights)
#   
#   temp2 <- matrix(1,nrow = nrow(data),ncol = ncol(data))
#   temp3 <- temp2 * heights
#   temp4 <- t( t( temp2 ) * widths )
#   temp2 <- matrix( "" , ncol = res , nrow = res )
#   
#   
#   for(i in 1:nrow(temp)){
#     for(j in 1:ncol(temp)){
#       temp2 <-
#     }
#   }
#   temp <- apply(temp,1,function(x){rep( x , widths/ )})
#   return(   )
# }
#
# m <- layout_matrix(c(1,2,3,4),2,2)
# m2 <- layout_matrix(c(1,2,3,"m"),2,2)

@@

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
taxa[is.na(taxa)] <- "NA"

par(mar=c(1,10,1,15))
x <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),2]
y <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),3]
# plot_sankey( data.frame( x[y!="NA"] , y[y!="NA"]), color_by = 1,
#        plot_labels = T, plot_weights = F,order_1_by = "total",
#        order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
#        pal=hue_pal()(length(unique(x))))

# plot_sankey( data.frame( y[y!="NA"] , bac_communities[y!="NA"]), color_by = 1,
#      plot_labels = T, plot_weights = F,order_1_by = "total",
#      order_2_by = "total",use_w1 = T,use_w2 = T,gapv = .02,gap2v = 0,smoothing = .1,
#      pal=hue_pal()(length(unique(y))))

@@

design <- model.matrix(~group_annotation, data=metadata[sample_use,])

@@

lrt <- glmLRT(fit,coef=2:8)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
colnames(top) <- sub("group_annotation","",colnames(top))
top <- cbind(logFC.Gard=0,top)

@@

image( t(top_picrust_data[h$order,][nrow(top_picrust_data):1,order(group_annotation[sample_use])]),
       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
       breaks = seq(-5,5,length.out = 100),axes=F)
points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
        col=pal[group_annotation[sample_use][ order(group_annotation[sample_use])]],
        pch=16)
points( seq(0,1,length.out = ncol(top_picrust_data)),rep(par("usr")[3],ncol(top_picrust_data))-.005,
        col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][order(group_annotation[sample_use])],
        pch=16,xpd=T)
points( seq(0,1,length.out = ncol(top_picrust_data)),rep(par("usr")[3],ncol(top_picrust_data))-.01,
        col= c("grey","red")[as.numeric(metadata$HIVstatus)][order(group_annotation[sample_use])],
        pch=16,xpd=T)

@@

mypar(2,3,mar=c(2,2,2,6))
set.seed(1)
UMAP_picrust <- uwot::umap(t(named_picrust_data),n_neighbors = 20,
                 metric = "correlation",min_dist = .2,spread = .4,
                 repulsion_strength = .4,negative_sample_rate = 3)
plot(UMAP_picrust,bg=pal[group_annotation[sample_use]],main="UMAP on top functional profile",
     pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
legend(par("usr")[2],par("usr")[4],title.adj = 0,
       legend = levels(factor(group_annotation)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
```

@@

```{r, echo = FALSE, eval = TRUE}
taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
dim(picrust_taxa)
dim(picrust_KO_seq)

@@

picrust_KO_seq <- rowsum(picrust_KO_seq , picrust_taxa[ match( rownames(picrust_KO_seq) , as.character(picrust_taxa$Seq_ID)  ),"final_name"])
picrust_KO_seq <- as.matrix(picrust_KO_seq)

@@

@@

## Figure 3

@@

TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
TRX_counts <- round(2^TRX - 1)
TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
sample_use <- colnames(TRX_counts)[colSums(TRX_counts)!=0]
TRX_counts <- TRX_counts[,sample_use]

library(edgeR)

design <- model.matrix(~  group_annotation, data=metadata[sample_use,])

lrt <- glmLRT(fit,coef=2:8)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
colnames(top) <- sub("group_annotation","",colnames(top))
top <- cbind(logFC.Gard=0,top)

saveRDS( y , "../../../results/TRX_EdgeR_estimations.rds" )
write.csv2( top , "../../../results/TRX_EdgeR_top_dge.csv" )

top_dge <- (top$PValue < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
top_dge <- rownames(top)[ top_dge ]

@@

h <- hclust( as.dist( (1- cor(t(top_TRX_counts)))/2 ), method = "ward.D2")
cuttoff <- 2
gene_module <- cutree(h, h=cuttoff)

layout(matrix(c(1,2,3,4,
              1,2,3,4),
              nrow = 2,ncol = 4,byrow = T),widths = c(1,2,.5,6))
par(mar=c(4,.1,2,.1))
plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.2),horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T,xlab="height")
# rect.dendrogram(rev(as.dendrogram(h)),h = 2,horiz = TRUE)
abline(v=cuttoff,xpd=F,col="red",lty=2)
points( rep(-.1,length(gene_module)),
        seq(length(gene_module),1,length.out = length(gene_module)),
        col=taxa_pal[factor(gene_module[h$order])],
        pch=16,cex=.5,xpd=T)

>>>>

---
title: "Integrated multi-omics analysis reveals Lactobacillus anti-inflammatory process in vaginal tissue"
subtitle: A demonstration of Rmarkdown using Herman Bumpus' data
author: |
  Author One^1^,
  Author Two^2^,
  Author Three^1,2^
keywords: "pandoc, r markdown, knitr"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  bookdown::pdf_document2:
      keep_tex: true
      toc: false
  bookdown::html_document2: default
  bookdown::word_document2: default
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[L]{XXXXXXXXXXX et al.}
- \fancyhead[R]{MANUSCRIPT SHORT TITLE}
- \usepackage{lineno}
- \linenumbers
linestretch: 1.5
link-citations: yes
linkcolor: blue
csl: nature.csl
cslreferences: nature.csl
bibliography: refs.bib
editor_options:
  chunk_output_type: console #inline
#params:
  #taxa_level: "Genus_lacto"
---
\footnotetext[1]{University of Nowhere}
\footnotetext[2]{University of Somewhere}
\footnotetext[3]{University of Lalaland}

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
params <- "Genus_lacto"
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(out.width = '100%',
                      dpi=600,
                      fig.width=10,
                      echo = FALSE,
                      # fig.pos = 'p', # Places figures on their own pages
                      eval = TRUE,
                      results = 'hide',
                      warning = FALSE,
                      message = FALSE )

# I usually load my libraries up front to keep things organized
library(bookdown)
library(knitr)
library(kableExtra)
library(dplyr)
library(tidyverse)
library(igraph)
library(phyloseq)
library(vegan)
library(ggpubr)
library(rafalib)
library(scales)
library(dendextend)
library(edgeR)
library(openxlsx)
remotes::install_github("czarnewski/niceRplots",force=T)
library(niceRplots)

@@

```{r, echo = FALSE, eval = TRUE}
# Load reasult from the analysis
metadata <- read.csv2("../../../results/metadata_integration.csv",row.names = 1)
TissueGroupM4 <- read.csv("../../../data/210411_TissueGroupM4.csv", row.names = 1)
#patient_groups <- factor(setNames(metadata$joint_clustering,metadata$ID))
patient_groups <- factor(setNames(metadata$ModifiedM4,metadata$ID))
bac_communities <- as.matrix(read.csv2("../../../results/bacterial_communities.csv",row.names = 1))[,1]

@@

```{r}
#color palettes
Set1 <- c("#4DAF4A", "#377EB8", "#E41A1C", "#FF7F00", "#984EA3", "#FFFF33", "#A65628", "#F781BF", "#999999")
pal <- c(Set1,RColorBrewer::brewer.pal(8,"Set2"),
         RColorBrewer::brewer.pal(9,"Pastel1"),RColorBrewer::brewer.pal(8,"Pastel2")) #color pallete for plots
heat_pal <- c("#000000", colorRampPalette(c("#000000","grey5","grey30","orange3","yellow","yellow","white"))(90))
cor_pal <- colorRampPalette(c("navy","white","firebrick"))(90)
```

```{r, echo = FALSE, eval = TRUE}
#Assign names to patient groups
# big_groups <- c(
#   g1_L.crispatus= "1 NonIners",
#   g2_L.iners=    "2 Iners",
#   g3_Gard=   "3 Gard",
#   g4_Prev=        "4 Prev",
#   g5_Oth=   "5 Other"
# )
big_groups <- c(
  g1_L.crispatus= "NonIners",
  g2_L.iners=     "Iners",
  g3_Gard=        "Gard",
  g4_Prev=        "Prev",
  g5_Oth=         "Oth"
)
# big_groups <- c(
#   g2_L.iners=    "5",
#   g3_L.iners2=   "4",
#   g1_L.crispatus= "2",
#   g5_Gard=        "3",
#   g6_Gard.Prev=   "6",
#   g7_Gard.Prev2=  "8",
#   g4_Gard.Lacto=  "7",
#   g8_Patho=       "1"
# )

group_annotation <- as.character(patient_groups)
group_annotation <- names(big_groups)[match(group_annotation,big_groups)]
group_annotation <- setNames(group_annotation,names(patient_groups))
#group_annotation <- as.factor(setNames(group_annotation,names(patient_groups)))

metadata$group_annotation <- group_annotation
#write.csv2(metadata,"../../../results/metadata_integration_.csv",row.names = T)

# modifications nessecary for dge analyser:
metadata <- metadata %>%
  mutate(ModifiedM4 = as.character(ModifiedM4_)) %>%
  mutate(sexwork_months=ifelse(is.na(.$sexwork_months),
                               mean(metadata$sexwork_months,na.rm = TRUE),.$sexwork_months))%>%
  mutate(Contraception=ifelse(.$Contraception=="no HC", "no_HC","DMPA")) %>%
  cbind(TissueGroupM4) %>%
  #as_data_frame() %>%
  #left_join(TissueGroupM4, by="ID") %>%
  #mutate(TissueGroupM4= factor(.$TissueGroupM4)) %>%
  select(ID, PatID, HIVstatus, Contraception, age, sexwork_months, everything())

#write.csv2(metadata,"../../../results/metadata_integration_.csv",row.names = T)

@@

#Assign names to bacterial communities
BC_groups <- c(
  BC01= "BC01_Peptococcus",
  BC02= "BC02_Lactobacillus.spp",
  BC03= "BC03_Micrococcus",
  BC04= "BC04_Gardnerella",
  BC05= "BC05_Paracoccus",
  BC06= "BC06_Pseudomonas",
  BC07= "BC07_Prvotella",
  BC08= "BC08_Faecalibacterium",
  BC09= "BC09_Escherichia/Shigella"
)
# BC_groups <- c(
#   BC01= "BC01_Bifidobacterium",
#   BC02= "BC02_Mycoplasma",
#   BC03= "BC03_Lactobacillus.spp",
#   BC04= "BC04_Escherichia.sp",
#   BC05= "BC05_Other1",
#   BC06= "BC06_Other2",
#   BC07= "BC07_Corynebacterium.spp",
#   BC08= "BC08_Other3",
#   BC09= "BC09_Ureaplasma.sp",
#   BC10= "BC10_Campylobacter",
#   BC11= "BC11_Mobiluncus_BVAB",
#   BC12= "BC12_Faecalibacterium",
#   BC13= "BC13_Dialister",
#   BC14= "BC14_P.timonesis",
#   BC15= "BC15_T.scotoductus"
# )

@@

par(mar=c(0,0,0,6))
plot( g , vertex.label.cex=0.000001 , vertex.color = pal[factor(group_annotation)] ,
      edge.width=  ( E(g)$weight / max(E(g)$weight)) ,
      vertex.size=10,main="", layout=l,
      edge.color=colorRampPalette(c("grey95","black"))(90) [ round( E(g)$weight / max(E(g)$weight) * 88 )+1 ] )
legend(par("usr")[2],par("usr")[4],title.adj = 0,
       legend = levels(factor(group_annotation)),
       xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)

@@

#######################
# BACTERIAL ABUNDANCE #
#######################
# for Genus_lacto taxonomy:
taxa <- read_csv("../../../results/SeqID_to_gen-lacto_tax-oth.csv") %>%
  select(-Sequence, -Seq_ID, -taxa_other, -Species) %>%
  unique() %>%
  filter(!(is.na(.$Genus_lacto))) %>%
  filter(!(.$Genus == "Granulicatella" & .$Family == "Enterococcaceae")) %>%
  column_to_rownames(., var = "Genus_lacto")
# For other_taxa taxonomy:
#taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
par(mar=c(4,5,1,12))
temp <- t(t(2^x-1)/colSums(2^x-1))

taxa[is.na(taxa)] <- "NA"
taxa <- taxa[rownames(temp),]
temp <- rowsum(temp , taxa$Class)
temp[is.nan(temp)] <- 0
o <- order( rowSums(temp) , decreasing = T)
sample_o <- order( group_annotation )
temp <- temp[o,sample_o]*100
#temp <- temp[-grep("NA",rownames(temp)),]
temp <- rbind(temp[1:16,], Other = colSums(temp[17:nrow(temp),]) )

taxa_pal <- c(RColorBrewer::brewer.pal(8,"Pastel2"),RColorBrewer::brewer.pal(8,"Pastel1"),"grey90")
barplot( temp, las=2 , border = NA , yaxs="i", xaxs="i",ylab="abundance",
         col = taxa_pal[factor(rownames(temp),levels = rownames(temp))] ,names.arg = rep("",ncol(temp)))
legend(x = par("usr")[c(2)],
       y = par("usr")[c(4)],title.adj = 0,
       legend = rownames(temp),xjust = 0,yjust = 1,
       bty = "n",pch = 22,pt.cex = 2,cex = 1,xpd=T,
       pt.bg = taxa_pal)
#
# points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-1,
#         col= c("grey","red")[(!is.na(metadata$RepeatedPCR))+1][sample_o],
#         pch=16,cex=.5,xpd=T)
# points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
#         col= c("grey","red")[as.numeric(metadata$HIVstatus)][sample_o],
#         pch=16,cex=.5,xpd=T)
  points( (1:ncol(temp))*1.2 -.5, rep(par("usr")[3],ncol(temp))-2,
        col= c("red","orange","grey")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation )],
        pch=15,cex=.8,xpd=T)
  text(par("usr")[1],par("usr")[3]-2,labels = "BV Diagnosis",pos=2,cex=.8,xpd=T)

@@

```{r, echo = FALSE, eval = TRUE,fig.asp=.6}
# Define Layout and set labels
layout(matrix(c(2,3,4,5,      6,7,11,11,          11,11,11,11,
                1,1,1,1,      1,8,11,11,          11,11,11,11,
                1,1,1,1,      1,9,12,12,         12,12,12,12,
                1,1,1,1,      1,10,13,13,         13,13,13,13,
                0,0,0,0,      0,0,14,14,         14,14,14,14 
                ) , 
              ncol= 12,byrow = T),widths = 1,heights = c(1.5,1.2,1.2,1.2,1.2))
# layout.show(3)
figlabels <- letters
#
# m <- matrix(as.character(c(1,2,3,4)),nrow = 2,ncol = 2)
# m
# m2 <- matrix(as.character(c(1,2,3,"m")),nrow = 2,ncol = 2)
# m2
#
# matrix(c(1,2,j,3),nrow = 2,ncol = 2)
# layout_matrix <- function(data, nrow, ncol){
#   return( matrix(as.character(data), nrow = nrow, ncol = ncol)  )
# }
# layout_matrix2 <- function(data, nrow, ncol, widths=c(1,2), heights=c(2,3), res=res){
#   temp <- matrix(as.character(data), nrow = nrow, ncol = ncol)
#   widths <- rep(widths,ncol(temp))[1:ncol(temp)]
#   heights <- rep(heights,nrow(temp))[1:nrow(temp)]
#   
#   cumsum(widths)*100/sum(widths)
#   cumsum(heights)*100/sum(heights)
#   
#   temp2 <- matrix(1,nrow = nrow(data),ncol = ncol(data))
#   temp3 <- temp2 * heights
#   temp4 <- t( t( temp2 ) * widths )
#   temp2 <- matrix( "" , ncol = res , nrow = res )
#   
#   
#   for(i in 1:nrow(temp)){
#     for(j in 1:ncol(temp)){
#       temp2 <-
#     }
#   }
#   temp <- apply(temp,1,function(x){rep( x , widths/ )})
#   return(   )
# }
#
# m <- layout_matrix(c(1,2,3,4),2,2)
# m2 <- layout_matrix(c(1,2,3,"m"),2,2)

@@

###################################
# BACTERIAL VAGINOSIS COMPARISSON #
###################################
# for Genus_lacto taxonomy:
taxa <- read_csv("../../../results/SeqID_to_gen-lacto_tax-oth.csv") %>%
  select(-Sequence, -Seq_ID, -taxa_other, -Species) %>%
  unique() %>%
  filter(!(is.na(.$Genus_lacto))) %>%
  filter(!(.$Genus == "Granulicatella" & .$Family == "Enterococcaceae")) %>%
  column_to_rownames(., var = "Genus_lacto")
# for other taxa
# taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
# taxa[is.na(taxa)] <- "NA"

#par(mar=c(1,10,1,15))
par(mar=c(1,10,1,15))
x <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),2]
y <- taxa[match(rownames(SNN_bacteria),rownames(taxa)),3]

@@

design <- model.matrix(~ group_annotation+Contraception+HIVstatus+age+sexwork_months, 
                       data=metadata[sample_use,])

@@

lrt <- glmLRT(fit,coef=2:9)
top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
colnames(top) <- sub("group_annotation","",colnames(top))
top <- cbind(LogFC.intercept=0,top) # with intercept

@@

image( t(top_picrust_data[h$order,][nrow(top_picrust_data):1,order(group_annotation[sample_use])]),
       col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
       breaks = seq(-5,5,length.out = 100),axes=F)
points( seq(0,1,length.out = length(sample_use)),rep(par("usr")[4],length(sample_use)),
        col=pal[factor(group_annotation[sample_use])[ order(group_annotation[sample_use])]],
        pch=16)
points( seq(0,1,length.out = ncol(top_picrust_data)),rep(par("usr")[3],ncol(top_picrust_data))-.005,
        col= c("red","orange","grey")[as.numeric(metadata$BV_Diagnosis_v3)][order( group_annotation )],
        pch=16,xpd=T)
points( seq(0,1,length.out = ncol(top_picrust_data)),rep(par("usr")[3],ncol(top_picrust_data))-.01,
        col= c("grey","red")[as.numeric(metadata$HIVstatus)][order(group_annotation[sample_use])],
        pch=16,xpd=T)

@@

mypar(2.5,2.5,mar=c(2.5,2.5,2,6))
set.seed(1)
UMAP_picrust <- uwot::umap(t(named_picrust_data),n_neighbors = 20,
                 metric = "correlation",min_dist = .2,spread = .4,
                 repulsion_strength = .4,negative_sample_rate = 3)
plot(UMAP_picrust,bg=pal[factor(group_annotation[sample_use])],main="UMAP on top functional profile",
     pch=21,xlab="UMAP1",ylab="UMAP2",frame=F,axes=F)
legend(par("usr")[2],par("usr")[4],title.adj = 0,
       legend = levels(factor(group_annotation)),xjust = 0,yjust = 1,
       bty = "n",pch = 21,pt.bg = pal,pt.cex = 1,xpd=T)
```

@@

```{r, echo = FALSE, eval = TRUE}
genus_lacto.fun <- function(df){
  genus_lacto <- df %>%
    mutate(Genus_lacto = .$Genus, .after=Seq_ID) %>% 
    mutate(Genus_lacto = ifelse(grepl("crispatus",.$Species), paste(.$Genus_lacto, "crispatus/acidophilus"), 
                                ifelse(grepl("reuteri",.$Species), paste(.$Genus_lacto, "reuteri/oris/frumenti/antri"),
                                  ifelse(grepl("gasseri",.$Species), paste(.$Genus_lacto,"gasseri/johnsonii/taiwanensis"),
                                    ifelse(grepl("murinus",.$Species), paste(.$Genus_lacto,"murinus/animalis/apodemi/salivarius"),
                                      ifelse(grepl("plantarum",.$Species), 
                                           paste(.$Genus_lacto, "plantarum/fabifermentans/composti/paraplantarum/graminis/fuchuensis"),
                                        ifelse(grepl("Lactobacillus",.$Genus), paste(.$Genus, .$Species), 
                                              as.character(.$Genus_lacto))))))) )
  return(genus_lacto)
}

picrust_taxa <- read_csv("../../../results/SeqID_to_gen-lacto_tax-oth.csv") %>%
  column_to_rownames(var = "Seq_ID") %>%
  mutate(Genus_lacto = ifelse(is.na(.$Genus_lacto), "NA", .$Genus_lacto))
  #mutate(Genus_lacto = ifelse(is.na(.$Genus_lacto), "NA_", .$Genus_lacto))

# for Genus_lacto taxonomy:
taxa <- read_csv("../../../results/SeqID_to_gen-lacto_tax-oth.csv") %>%
  select(-Sequence, -Seq_ID, -taxa_other, -Species) %>%
  unique() %>%
  filter(!(is.na(.$Genus_lacto))) %>%
  filter(!(.$Genus == "Granulicatella" & .$Family == "Enterococcaceae")) %>%
  column_to_rownames(., var = "Genus_lacto")
# for other taxa
#taxa <- read.csv2("../../../results/taxonomy.csv",row.names = 1,stringsAsFactors = F,as.is = T)
dim(picrust_taxa)
dim(picrust_KO_seq)

@@

picrust_KO_seq <- rowsum(picrust_KO_seq , picrust_taxa[ match( rownames(picrust_KO_seq) , rownames(picrust_taxa)  ), params])
picrust_KO_seq <- as.matrix(picrust_KO_seq)

@@

#p <- sub(" [(].*","",names(pvalues))
#x <- p

@@

## Figure 3
## individual group comparison DGE
```{r}
TRX <- datasets_all_samples[["Tissue_RNAseq_V3_normalized"]]
TRX_counts <- round(2^TRX - 1)
TRX_counts <- TRX_counts [ rowSums(TRX_counts>0)>= 1 , ]
sample_use <- colnames(TRX_counts)[colSums(TRX_counts)!=0]
TRX_counts <- TRX_counts[,sample_use]

library(edgeR)
#group <- "TissueGroupM4"
#group <- "group_annotation"
edgR_dge.fun <- function(group, confound) {
  groups <- metadata[sample_use, group]

  if(confound==FALSE) {
    c <- "_no_conf"
    design <- model.matrix(~  groups, data=metadata[sample_use,])} # without confounders
  else{
    c <- "_w_conf"
    # design <- model.matrix(~groups+HIVstatus+Contraception+age+sexwork_months,
    #                      data=metadata[sample_use,])
    design <- model.matrix(~groups+HIVstatus+Contraception,
                         data=metadata[sample_use,])
  }
  #design <- model.matrix(~0 + groups+Contraception+HIVstatus+age+sexwork_months, 
                         #data=metadata[sample_use,]) # without intercept

  colnames(design) <- sub("groups","",colnames(design))
  colnames(design) <- sub("(Intercept)","Intercept",colnames(design))

  #lrt <- glmLRT(fit,coef=1:ncol(design)) # without intercept
  lrt <- glmLRT(fit,coef=2:5) # with intercept
  top <- topTags(lrt,adjust.method = "BH",n = "all",sort.by = "p.value")[[1]]
  colnames(top) <- sub(group,"",colnames(top))
  top <- cbind(LogFC.intercept=0,top) # with intercept
  top <-  rownames_to_column(top, var = "Genes")

  saveRDS( y , paste0("../../../results/TRX_EdgeR_estimations_",group,c,".rds" ))
  #write.xlsx(top, file=paste0("../../../results/DGE/","dge_heatmap_",group,c,".xlsx"))
  write.xlsx(top, file=paste0("../../../results/DGE/","dge_heatmap_",group,"_2:5_conf_2.xlsx"))
  #write.csv2( top , paste0("../../../results/DGE/TRX_EdgeR_top_dge_",group,".csv" ))

return(list(design=design, y=y, fit=fit, lrt=lrt, top=top, c=c))
}

dge_M4 <- edgR_dge.fun("ModifiedM4", confound=T)
#dge_M4_nonconfound <- edgR_dge.fun("ModifiedM4", confound=F)
#dge_TissueM4 <- edgR_dge.fun("TissueGroupM4", confound=T)

@@

# group_name <- "ModifiedM4"
# top <- dge_M4$top
# y <- dge_M4$y
# c <- dge_M4$c
gene_mod_enrich.fun <- function(top, y, c, db, group_name, cuttoff){
  p_val <- pull(top, "PValue")
  top_dge <- (p_val < 0.01) #& (rowSums( abs(top[,grep("logFC",colnames(top))]) >= log2(1.5) ) >= 1)
  top_dge <- top$Genes[ top_dge ]

@@

  h <- hclust( as.dist( (1- cor(t(top_TRX_counts)))/2 ), method = "ward.D2")
  cuttoff <- cuttoff[names(cuttoff) == db]
  gene_module <- cutree(h, h=cuttoff)
